Generalised gravitational burst waveform generation with Generative Adversarial Networks

### 2020-04-06 Paper_draft/main.tex
update

\title{Generative Adversarial Networks and Gravitational Waves}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/train.pdf}
    \caption{Examples of burst training signals. From left to right: sine-gaussian, ringdown, white-noise burst, Gaussian, binary-black hole merger.}
    \label{fig:train}
\end{figure}

\subsection{Applying Antenna responses}
We consider a two detector case where the generator should be trained to output two identical signals with a physical time shift representing the time of flight between detectors and the suppression of amplitudes due to interaction with the detector. To achieve this, the generator is trained to output a single waveform that is then put through a non-trainable ``response" layer as the final layer of the generator. This layer creates a copy of its input, shifts it along in time and applies antenna responses to both the input and shifted signal. The result is a 2x1024 times series waveform representing an overlay data stream from two detector outputs. An example of each signal family is shown in Fig. \ref{fig:train}.

\subsection{Architecture details}
We base the architecture on (cite DCGANS) and adopt the suggestions of (cite wavegan paper) lengthening one-dimensional convolution kernels on both the generator and discriminator. The class label is interpreted as an additional channel early in the generator model. This is achieved by projecting the class inputs as a learned embedding layer, into a fully connected layer which can then be concatenated channel-wise to $\textbf{z}$. The rest of the model is fully convolutional, upsampled using strided transposed convolutions with batchnormalisation in the first layer and ReLU activations throughout with the exception of Tanh for the output layer. Each transposed convolutional layer uses a kernel size of 18x1 and stride 2.
The discriminator network mirrors that of the generator without batch normalization, using LeakyReLU activations, SpatialDropout, and a 2-stride convolution for downsampling. The main difference is that the discriminator has two output layers: the first output is a single node activated by a Sigmoid that can be interpreted as the realness of the the signal, the second output is 5 nodes activated using the softmax function predicting the class of the input. This model is trained with binary cross entropy for the first output and sparse categorical cross-entropy for the second output.
Neural networks and subsequently GANs have multiple parameters a developer can tune when designing the model referred to as hyperparameters. The final network design used in this work comes from the use of trial and error and the initial designs influenced by the available literature. After tuning the multiple hyperparamters (Table \ref{Tab:hyperparameters}), the GAN is trained for 600000 iterations which in real time is $\Or$(1) day.
\begin{comment}
include Plot_NN to show generator architecture
\end{comment}

@@

\subsection{Waveform quality}
The generator network is a function G : $\mathbf{z},\mathbf{c},\mathbb{\textbf{s}}$ $\in$ $\mathbb{R}^{100}$ $\to$ $\mathbb{R}^{1024\times2}$, where $\mathbf{z},\mathbf{c},\mathbb{\textbf{s}}$ are the latent vector, class vector and sky positions respectively. Given a latent vector randomly sampled from a normal distribution with zero mean and unit variance, a class label which is represented by a 120 dimensional vector for each class and antenna responses, the results from the generator can be seen in Fig. \ref{fig:gen_signals}. Each plot shows the output of the generator after given randomised $z$, $s$ and one of the five class vectors.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/gen_sigs.png}
    \caption{Generated waveforms after training and conditioning on five classes. The generator is able to capture the characteristics of each waveform and structure the class space to give control over which waveform to generate. Depending on the orientation of the detector with respect to a hypothetical signal in the sky, the waveforms may appear inverted, shifted in time and their strain attenuated.}
    \label{fig:gen_signals}
\end{figure}
\subsection{Latent space interpolation}
In this section we explore the latent space formed by the generator by interpolation. Keeping the class space fixed and interpolating between to points in z we can see the full effect shown in Fig. \ref{fig:z_interp}. We can see that each plot shows plausible waveforms suggesting that the generator has constructed a smooth space unlike the discrete training case. Additionally as each class is given the same latent points to interpolate over, we can see that the waveforms cluster together with respect to their parameters. Visually, the sing-gaussian and ring-down waveforms share similar frequencies and the other signals show similar decays and starting epochs. The only exception is BBH waveforms, which is expected as they were trained with more variety of parameters and coinsitently have their peaks in the last quarter of the time series.
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/fixed_class_slerp.png}
    \caption{Interpolations between two random latent points in z. Moving right along the columns moves a step in z while holding c fixed. Each row changes the class vector while maintaining the same interpolations in z.}
    \label{fig:z_interp}
\end{figure}
\subsection{Class space interpolation}
In order to explore the class space we keep the latent vector held constant and interpolate through the 5 classes. We construct a path between the 5 waveforms and show that the space is populated enough to allow for smooth transitions between classes. 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/fixed_z_slerp.png}
    \caption{Class space interpolation with latent space held constant throughout.}
    \label{fig:c_interp}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{itemize}
\item Summarise the paper
\item Dedicate a paragraph to each of the key results discussed in the previous
section
\item Have at least one paragraph on the future directions of this work
\item Conclude with a positive paragrpah about the potential uses and impact of
the approach.
\end{itemize}

\section*{References}
\bibliography{iopart-num}

>>>>

\title{Generalised gravitational burst waveform generation with Generative Adversarial Networks}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.7]{figures/training.pdf}
    \caption{Examples of burst training signals. From left to right: sine-gaussian, ringdown, white-noise burst, Gaussian, binary-black hole merger.}
    \label{fig:train}
\end{figure}

\subsection{Applying Antenna responses}
We consider a two detector case in which the generator is trained to output two identical signals with a physical time shift representing the time of flight between detectors and the suppression of amplitudes due to interaction with the detector. To achieve this, the generator is trained to output a single waveform that is then put through a non-trainable ``response" layer as the final layer of the generator. This layer creates a copy of its input, shifts it along in time and applies antenna responses to both the input and shifted signal. The result is a 2x1024 times series waveform representing an overlay data stream from two detector outputs. An example of each signal family is shown in Fig. \ref{fig:train}.

\subsection{Architecture details}
We base the architecture on deep convolutional generative adversarial networks \cite{Radford2015} and adopt the suggestions of \cite{DBLP:journals/corr/abs-1809-11096} lengthening one-dimensional convolution kernels on both the generator and discriminator. The class label is interpreted as an additional channel early in the generator model. This is achieved by projecting the class inputs as a learned embedding layer, into a fully connected layer which can then be concatenated channel-wise to $\textbf{z}$. The rest of the model is fully convolutional, upsampled using strided transposed convolutions with batchnormalisation in the first layer and ReLU activations throughout with the exception of Tanh for the output layer. Each transposed convolutional layer uses a kernel size of 18x1 and stride 2.
The discriminator network mirrors that of the generator without batch normalization, using LeakyReLU activations, SpatialDropout, and a 2-stride convolution for downsampling. The main difference is that the discriminator has two output layers: the first output is a single node activated by a Sigmoid that can be interpreted as the realness of the the signal, the second output is 5 nodes activated using the softmax function predicting the class of the input. This model is trained with binary cross entropy for the first output and sparse categorical cross-entropy for the second output.
Neural networks and subsequently GANs have multiple parameters a developer can tune when designing the model referred to as hyperparameters. The final network design used in this work comes from the use of trial and error and the initial designs influenced by the available literature. After tuning the multiple hyperparamters (Table \ref{Tab:hyperparameters}), the GAN is trained for 600000 iterations which in real time is $\Or$(1) day.
\begin{comment}
include Plot_NN to show generator architecture
\end{comment}

@@

\subsection{Waveform quality}
The generator network is a function G : $\mathbf{z},\mathbf{c},\mathbb{\textbf{s}}$ $\in$ $\mathbb{R}^{100}$ $\to$ $\mathbb{R}^{1024\times2}$, where $\mathbf{z},\mathbf{c},\mathbb{\textbf{s}}$ are the latent vector, class vector and sky positions respectively. Given a latent vector randomly sampled from a normal distribution with zero mean and unit variance, a class label which is represented by a 120 dimensional vector for each class and antenna responses, the results from the generator can be seen in Fig. \ref{fig:gen_signals}. Depending on the orientation of the detector with respect to a hypothetical signal in the sky, the waveforms may appear inverted, shifted in time and their strain attenuated. Each plot shows the output of the generator after given randomised $z$, $s$ and one of the five class vectors.

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/generations.pdf}
    \caption{Generated waveforms after training and conditioning on five classes. The generator will output 2 identical waveforms shifted in time (shown by light and dark blue lines) and their amplitudes reduced by detector response. The generator is able to capture the characteristics of each waveform and structure the class space to give control over which waveform to generate. Each row shows a random assortment of one of the five classes the GAN is trained on. }
    \label{fig:gen_signals}
\end{figure}

\subsection{Latent space interpolation}
In this section we explore the latent space formed by the generator by interpolation. Keeping the class space fixed and interpolating between to points in z we can see the full effect shown in Fig. \ref{fig:z_interp}. We can see that each plot shows plausible waveforms suggesting that the generator has constructed a smooth space unlike the discrete training case. Additionally as each class is given the same latent points to interpolate over, we can see that the waveforms cluster together with respect to their parameters. Visually, the sing-gaussian and ring-down waveforms share similar frequencies and the other signals show similar decays and starting epochs. The only exception is BBH waveforms, which is expected as they were trained with more variety of parameters and coinsitently have their peaks in the last quarter of the time series.
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/fix_c_interp_z.pdf}
    \caption{Interpolations between two random latent points in z. Each row uniformly interpolates between two points in $\mathbf{z}$ keeping the class fixed. Each subplot shows the next interpolated signal (dark line) and the fixed point in $\mathbf{z}$ (light line). }
    \label{fig:z_interp}
\end{figure}
\subsection{Class space interpolation}
In order to explore the class space we keep the latent vector held constant and interpolate through the 5 classes. We construct a path between the 5 waveforms and show that the space is populated enough to allow for smooth transitions between classes. 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/fix_z_interp_c.pdf}
    \caption{Class space interpolation with latent space held constant throughout. }
    \label{fig:c_interp}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/4_corners.pdf}
    \caption{We generate the four corners of the grid, then interpolate between both the z and c vectors. Although the network was only trained on binary attributes for c, interpolation results show that the computed latent space is smooth between attributes, and not simply discrete to those shown in the training set.}
    \label{fig:4_c_interp}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}
\begin{itemize}
\item Summarise the paper
\item Dedicate a paragraph to each of the key results discussed in the previous
section
\item Have at least one paragraph on the future directions of this work
\item Conclude with a positive paragrpah about the potential uses and impact of
the approach.
\end{itemize}
\end{comment}

\newpage
\section*{References}
\bibliography{iopart-num}
